\(ler\) : LVK (LIGO-Virgo-KAGRA collaboration) event (compact-binary mergers) rate calculator and simulator

### 2024-07-02 paper/paper.tex
mac

      \affil[1]{Department of Physics, The Chinese University of Hong
Kong, Shatin, New Territories, Hong Kong}
      \affil[2]{The Inter-University Centre for Astronomy and
Astrophysics (IUCAA), Post Bag 4, Ganeshkhind, Pune 411007, India}
      \affil[3]{Kavli Institute for the Physics and Mathematics of the
Universe (IPMU), 5-1-5 Kashiwanoha, Kashiwa-shi, Chiba 277-8583, Japan}
      \affil[4]{Department of Physics, Utrecht University,
Princetonplein 1, 3584 CC Utrecht, The Netherlands}
      \affil[5]{Nikhef -- National Institute for Subatomic Physics,
Science Park, 1098 XG Amsterdam, The Netherlands}
      \affil[6]{Center for Cosmology, Particle Physics and Phenomenology
- CP3, Universit\textquotesingle e Catholique de Louvain,
Louvain-La-Neuve, B-1348, Belgium}
      \affil[7]{Royal Observatory of Belgium, Avenue Circulaire, 3, 1180
Uccle, Belgium}
      \affil[8]{Departament de F\textquotesingle isica, Universitat de
les Illes Balears, IAC3-IEEC, Crta. Valldemossa km 7.5, E-07122 Palma,
Spain}
  \date{\vspace{-7ex}}

@@

Gravitational waves (GWs) are ripples in the fabric of space and time
caused by the acceleration of unevenly distributed mass or masses.
Observable GWs are created especially during the violent cosmic events
of merging compact binaries, such as `binary black holes' (BBH), `binary
neutron stars' (BNS), and `neutron star and black hole pair' (NSBH). GWs
emitted by these events can be distorted or magnified by the
gravitational fields of massive objects such as galaxies or galaxy
clusters, a phenomenon known as gravitational lensing. Profound
comprehension of gravitational lensing's impact on GW signals is
imperative to their accurate interpretation and the extraction of
astrophysical insights therein. For this purpose, statistical modelling
of GWs lensing can provide valuable insights into the properties of the
lensing objects and GW sources. Such statistics require accurate and
efficient means to calculate the detectable lensing rates, which depend
on up-to-date modeling and implementation of lens and source properties
and their distribution. The outcomes of these computational analyses not
only contribute to generating dependable forecasts but also play an
important role in validating forthcoming lensing events (Janquart et al.
2023).

@@

\begin{itemize}
\tightlist
\item
  Sampling GW Source Properties:

  \begin{itemize}
  \tightlist
  \item
    For the unlensed events, The sampling distribution
    \((\,R_m(z_s)\,)\) for the source's redshift \((\,z_s\,)\) is
    derived from the merger rate density of compact binaries, which, in
    turn, is based on the star formation rate. The code is meticulously
    designed to enable the straightforward integration of future updates
    or user-specified distributions of these sources.
  \item
    Intrinsic and extrinsic parameters \((\theta)\) of GW sources are
    sampled using prior distributions \((P(\theta))\) from the gwcosmo
    (gwcosmo Contributors 2022) and bilby (Ashton et al. 2019) packages,
    with options for users to input custom distributions.
  \end{itemize}
\item
  Sampling of lens galaxies attributes and source red-shifts:

  \begin{itemize}
  \tightlist
  \item
    For lensed case, the source redshift \((z_s)\) is sampled under the
    strong lensing condition \((\text{SL})\), based on the precomputed
    probability of strong lensing with source at \(z_s\)
    \((\text{optical depth: }P\left(\text{SL}|z_s\right) \text{ or }\tau(z_s)\,)\).
    This probability can be recalculated for specified configurations of
    lens galaxies, leveraging \emph{multiprocessing} and \emph{njit}
    functionalities for enhanced efficiency.
  \item
    The package utilizes the Elliptical Power Law with external shear
    (EPL+Shear) model (Wempe et al. 2022) for galaxy parameters
    \((\theta_L)\) sampling, following Wierda et al. (2021). Rejection
    sampling is applied on the above samples on condition that whether
    the event is strongly lensed or not,
    \(P\left(\text{SL}|z_s,\theta_L\right)\).
  \end{itemize}
\item
  Generation of image properties:

  \begin{itemize}
  \tightlist
  \item
    Source position \((\beta)\) is sampled from the caustic in the
    source plane.
  \item
    Sampled lens properties and source position is fed in
    \emph{lenstronomy} (Birrer et al. 2021) to generate properties of
    the images. This is the slowest part of the entire simulation, which
    \(ler\) tackles through parallelization with multiprocessing.
  \item
    Image properties like magnification \((\mu_i)\) and time delay
    (\(\Delta t_i\)) modifies the original source signal strength,
    changing the signal-to-noise ratio SNR and our ability to detect.
  \end{itemize}
\item
  Calculation of Detectable Merger Rates Per Year:

@@

The authors express their sincere appreciation for the significant
contributions that have been instrumental in completing this research.
Special thanks are extended to the academic advisors for their
invaluable guidance and steadfast support. The collaborative efforts and
enriching discussions with research colleagues significantly enhanced
the study's quality. Acknowledgement is given to the Department of
Physics, The Chinese University of Hong Kong, for the Postgraduate
Studentship that facilitated this research. Further gratitude is
extended to the Netherlands Organisation for Scientific Research (NWO)
for their support. The authors also recognize the contributions of
individuals who added empirical depth to this work. Appreciation is
conveyed for the computational resources provided by the LIGO
Laboratory, supported by National Science Foundation Grants
No.~PHY-0757058 and No.~PHY-0823459.

>>>>

      \affil[1]{Department of Physics, The Chinese University of Hong
Kong, Shatin, New Territories, Hong Kong}
      \affil[2]{The Inter-University Centre for Astronomy and
Astrophysics (IUCAA), Post Bag 4, Ganeshkhind, Pune 411007, India}
      \affil[3]{Kavli Institute for the Physics and Mathematics of the
Universe (IPMU), 5-1-5 Kashiwanoha, Kashiwa-shi, Chiba 277-8583, Japan}
      \affil[4]{Department of Physics, Utrecht University,
Princetonplein 1, 3584 CC Utrecht, The Netherlands}
      \affil[5]{Nikhef -- National Institute for Subatomic Physics,
Science Park, 1098 XG Amsterdam, The Netherlands}
      \affil[6]{Center for Cosmology, Particle Physics and Phenomenology
- CP3, Universit\textquotesingle e Catholique de Louvain,
Louvain-La-Neuve, B-1348, Belgium}
      \affil[7]{Royal Observatory of Belgium, Avenue Circulaire, 3, 1180
Uccle, Belgium}
      \affil[8]{Departament de Física, Universitat de les Illes Balears,
IAC3-IEEC, Crta. Valldemossa km 7.5, E-07122 Palma, Spain}
  \date{\vspace{-7ex}}

@@

Gravitational waves (GWs) are ripples in the fabric of space and time
caused by the acceleration of unevenly distributed mass or masses.
Observable GWs are created especially during the violent cosmic events
of merging compact binaries, such as `binary black holes' (BBHs),
`binary neutron stars' (BNSs), and `neutron star and black hole pair'
(NSBHs). GWs emitted by these events can be distorted or magnified by
the gravitational fields of massive objects such as galaxies or galaxy
clusters, a phenomenon known as gravitational lensing. Profound
comprehension of gravitational lensing's impact on GW signals is
imperative to their accurate interpretation and the extraction of
astrophysical insights therein. For this purpose, statistical modelling
of GWs lensing can provide valuable insights into the properties of the
lensing objects and GW sources. Such statistics require accurate and
efficient means to calculate the detectable lensing rates, which depend
on up-to-date modeling and implementation of lens and source properties
and their distribution. The outcomes of these computational analyses not
only contribute to generating dependable forecasts but also play an
important role in validating forthcoming lensing events (Janquart et al.
2023).

@@

\begin{itemize}
\tightlist
\item
  Sampling GW source properties:

  \begin{itemize}
  \tightlist
  \item
    For the unlensed events, the sampling distribution
    \((\,R_m(z_s)\,)\) for the source's redshift \((\,z_s\,)\) is
    derived from the merger rate density of compact binaries, which, in
    turn, is based on the star formation rate. The code is meticulously
    designed to enable the straightforward integration of future updates
    or user-specified distributions of these sources.
  \item
    Intrinsic and extrinsic parameters \((\theta)\) of GW sources are
    sampled using prior distributions \((P(\theta))\) from the gwcosmo
    (gwcosmo Contributors 2022) and bilby (Ashton et al. 2019) packages,
    with options for users to input custom distributions.
  \end{itemize}
\item
  Sampling of lens galaxy attributes and source red-shifts:

  \begin{itemize}
  \tightlist
  \item
    For the lensed case, the source redshift \((z_s)\) is sampled under
    the strong lensing condition \((\text{SL})\), based on the
    precomputed probability of strong lensing with a source at \(z_s\)
    \((\text{optical depth: }P\left(\text{SL}|z_s\right) \text{ or }\tau(z_s)\,)\).
    This probability can be recalculated for specified configurations of
    lens galaxies, leveraging \emph{multiprocessing} and \emph{njit}
    functionalities for enhanced efficiency.
  \item
    The package utilizes the Elliptical Power Law with external shear
    (EPL+Shear) model (Wempe et al. 2022) for galaxy parameters
    \((\theta_L)\) sampling, following Wierda et al. (2021). Rejection
    sampling is applied on the above samples depending on whether the
    event is strongly lensed or not,
    \(P\left(\text{SL}|z_s,\theta_L\right)\).
  \end{itemize}
\item
  Generation of image properties:

  \begin{itemize}
  \tightlist
  \item
    The source position \((\beta)\) is sampled from the caustic in the
    source plane.
  \item
    Sampled lens properties and source positions are fed to
    \emph{lenstronomy} (Birrer et al. 2021) to generate properties of
    the images. This is the slowest part of the entire simulation, which
    \(ler\) tackles through parallelization with multiprocessing.
  \item
    Image properties like magnification \((\mu_i)\) and time delay
    (\(\Delta t_i\)) modify the original source signal strength,
    changing the signal-to-noise ratio (SNR) and our ability to detect.
  \end{itemize}
\item
  Calculation of detectable merger rates per year:

@@

The authors express their sincere appreciation for the significant
contributions that have been instrumental in completing this research.
Special thanks are extended to the academic advisors for their
invaluable guidance and steadfast support. The collaborative efforts and
enriching discussions with research colleagues significantly enhanced
the study's quality. Acknowledgement is given to the Department of
Physics, The Chinese University of Hong Kong, for the Postgraduate
Studentship that facilitated this research. Further gratitude is
extended to the Netherlands Organisation for Scientific Research (NWO)
for their support. The authors also recognize the contributions of
individuals who added empirical depth to this work. Appreciation is
conveyed for the computational resources provided by the LIGO
Laboratory, supported by National Science Foundation Grants
No.~PHY-0757058 and No.~PHY-0823459. N. Singh and D. Keitel are
supported by Universitat de les Illes Balears (UIB); the Spanish Agencia
Estatal de Investigación grants CNS2022-135440, PID2022-138626NB-I00,
RED2022-134204-E, RED2022-134411-T, funded by
MICIU/AEI/10.13039/501100011033, the European Union
NextGenerationEU/PRTR, and the ERDF/EU; and the Comunitat Autònoma de
les Illes Balears through the Direcció General de Recerca, Innovació I
Transformació Digital with funds from the Tourist Stay Tax Law
(PDR2020/11 - ITS2017-006) as well as through the Conselleria
d'Economia, Hisenda i Innovació with grant numbers SINCO2022/6719
(European Union NextGenerationEU/PRTR-C17.I1) and SINCO2022/18146
(co-financed by the European Union and FEDER Operational Program
2021-2027 of the Balearic Islands).
